### 2020-11-30 paper/localpsf.tex
editing intro

% REQUIRED
\begin{abstract}
  We present an efficient matrix-free method for approximating locally translation invariant operators with non-negative integral kernels. The idea of the method is to compute impulse responses of the operator at a collection of well-spaced points, then interpolate these impulse responses to approximate impulse responses at other arbitrary points. We convert the resulting approximation to hierarchical matrix format, which allows us to perform matrix inversion or other matrix operations using fast hierarchical matrix arithmetic. Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources, and special care is taken to ensure that the supports of the impulse responses within each batch do not overlap. We apply the method to approximate Hessians in large-scale PDE-constrained inverse problem with highly informative data. Numerical results demonstrate that our method substantially outperforms existing Hessian approximation methods which are based on low-rank approximation. Our method is able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
%  Within each Dirac comb, the point sources must be spaced sufficiently far apart so that the impulse responses to those point sources do not overlap. We estimate the required spacing of point sources by applying the operator to a small number of constant, linear, and quadratic functions, then postprocessing the results.
\end{abstract}

@@

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $a:L^2(\Omega) \times L^2(\Omega) \rightarrow \mathbb{R}$,
\begin{equation}
	\label{eq:kernel_representation}
	a(v,u) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
\end{equation}
be a bilinear form with non-negative integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$, $A(y,x) \ge 0$. Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying. That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is defined by
\begin{equation*}
	(\mathcal{A}u)(v) := a(v,u).
\end{equation*}
We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$. Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$. Integral kernel representation \eqref{eq:kernel_representation}, is assumed to exist,
%e.g., by the Schwarz kernel theorem,
but one cannot easily access entries $A(y,x)$ of the kernel. Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.

>>>>

% REQUIRED
\begin{abstract}
  We present an efficient matrix-free method for approximating locally translation invariant operators with locally supported non-negative integral kernels. The idea of the method is to compute impulse responses of the operator at a collection of well-spaced points, then interpolate these impulse responses to approximate impulse responses at other arbitrary points. We convert the resulting approximation to hierarchical matrix format, which allows us to perform matrix inversion or other matrix operations using fast hierarchical matrix arithmetic. Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources, and special care is taken to ensure that the supports of the impulse responses within each batch do not overlap. We apply the method to approximate Hessians in large-scale PDE-constrained inverse problem with highly informative data. Numerical results demonstrate that our method substantially outperforms existing Hessian approximation methods which are based on low-rank approximation. Our method is able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
%  Within each Dirac comb, the point sources must be spaced sufficiently far apart so that the impulse responses to those point sources do not overlap. We estimate the required spacing of point sources by applying the operator to a small number of constant, linear, and quadratic functions, then postprocessing the results.
\end{abstract}

@@

In this paper, we present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, with locally supported non-negative integral kernels. By ``matrix-free'', we mean that $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$. Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.

Specfically, we consider operators $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
\begin{equation*}
(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
\end{equation*}
where $\Omega \subset \mathbb{R}^d$ is a bounded domain, and the integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$ is non-negative in the sense that  $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$. 
%Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying. That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is 
%We approximate the operator $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ defined by
%\begin{equation*}
%	(\mathcal{A}u)(v) := a(v,u).
%\end{equation*}
%We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$. Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$. 
Integral kernel representation \eqref{eq:kernel_representation}, is assumed to exist,
%e.g., by the Schwarz kernel theorem,
but one cannot easily access entries $A(y,x)$ of the kernel.

We further assume that $A$ is locally supported in the sense that the impulse response function, 
\begin{equation*}
	\varphi_x(y) := A(y, x)
\end{equation*}
is supported in a region near $x$.
